---
title: "My Simple Logistic Regression"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<!-- Find some data of interest to you that can be analyzed with a logistic regression. Complete your Logistic Regression Analysis using this file. 

Hint, run the command 

> data() 

in your Console to see available datasets in R that you could use for this Analysis. You are also welcome to collect data yourself, or find other available datasets online for this Analysis. Just be careful to not use anyone else's "analysis" for this Analysis. Also, no conducting surveys unless you have IRB approval to do so. -->

```{r, include=FALSE, echo = FALSE}
library(mosaic)
library(ResourceSelection) 
library(readr)
library(tidyverse)
library(pander)

```

```{r import and clean the data, echo=FALSE, message=FALSE, warning=FALSE}
dat <- read_csv("/Users/andreasmartinson/Documents/School/1 - 2018 Fall Semester Classwork/MATH 325/csv_pwa/ss16pwa.csv")

dat <- dat %>% 
  select(AGEP,MAR) %>% 
  mutate(Age = as.numeric(AGEP))
```

####Background

This data was taken from the US Census Bureau data of Washington residents from 2012-2016. This is a rich dataset with many parameters, but I just wanted to focus on age and if the individual was married or not. The dataset can be found here: https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_pums_csv_2012_2016&prodType=document and is titled "Washington Population Records". I wanted to use the US records file, but since it was 2.3 GB it was going to take an hour to download, and therefore I chose to go with the Washington data subset, which still had over 350,000 rows of information. The married column (denoted as "MAR"), has 5 values, with the following meanings: 1 .Married
2 .Widowed
3 .Divorced
4 .Separated
5 .Never married or under 15 years old

By using the 5th number of "never married" as successful, or y = 1, and the other numbers as "failures"" (being married), I can perform a logistic regression on the data. 

####Question and Hypothesis

I wanted to figure out if I could predict the likelihood of a person being married based on their age. Therefore, my question is, "Can I fit a logistic model to predict the likelihood someone is married by using their age?"

The hypothesis is as follows:

$$
  H_0: \beta_1 = 0 \\
  H_a: \beta_1 \neq 0
$$

#### Logistic Model

The following model is used to help predict the likelihood of having married once at various ages:
$$
  P(Y_i = 1|x_i) = \frac{e^{\beta_0+\beta_1 x_i}}{1+e^{\beta_0 + \beta_1 x_i}} = \pi_i
$$
where for observation $i$: 

* $Y_i = 1$ is interpreted as not getting married, 
* $Y_i=0$ is interpreted as getting married at least once, and 
* $x_i$ is interpreted as the person's age. 

<br />

#### Fitting the Model

The following r code as used in order to perform the logistical regression:

```{r, echo = FALSE}
my_glm <- glm(MAR == 5 ~ Age, data = dat, family = binomial)
pander(summary(my_glm))
```

Thus the estimated model for $\pi_i$ is given by
$$
  P(Y_i = 1|x_i) \approx \frac{e^{3.982011-0.122036 x_i}}{1+e^{3.982011-0.122036 x_i}} = \hat{\pi}_i
$$
where $b_0 = 15.043$ is the value of the `(Intercept)` which estimates $\beta_0$ and $b_1 = -0.232$ is the value of `Age` which estimates $\beta_1$.

Tthe $p$-value for the test of `Age` shows a significant result $(p = <2e-16)$ giving sufficient evidence to conclude that $\beta_1 \neq 0$, this means that age does affect the likelihood of being married or not.

<br />

#### Visualizing the Model

The following is a plot of the data presented above.

```{r, echo = FALSE}
plot(MAR == 5 ~ Age, data = dat, xlab = "Age", ylab = "Probability of not being Married", main = "Logistic Regression of Marriage Data from 2012-2016", pch = 16)
curve(exp(3.982011-0.122036*x)/(1+exp(3.982011-0.122036*x)), add=TRUE)
```

This plot further illustrates the point that this data is a good dataset to perform the logistic regression on because there is a steep curve in the line, going all the way from 1 to 0 and forming a nice s-curve. 

<br />

#### Validating the data

I still need to do some diagnosis and check to see if this data really is appropriate for a logistic regression. Since there are many duplicate x-values, I will use the pchisq() function to check for goodness of fit below. 

```{r}
pander(pchisq(227463, 353679, lower.tail = FALSE))
```

The data shows that this model is a good fit for the data and the logistic regression is appropriate. 

#### Prediction 

I want to put the model to the test and see what the probability is that a 25 year old is not married. 

```{r, echo = FALSE}
pander(predict(my_glm, data.frame(Age = 25), type = 'response'))
```

The result is the there is a 71.73% chance that I am not married if I am a Washington resident. Also below, I did a quick calculation to see how quickly the chance of being not married decreases with age. 

Using the following equation, I can find how much the chances of not being married decreases with age: $e^{b_1} = e^{-0.122036} \approx 0.89$. This is computed below. 

```{r}
pander(exp(-0.122036))
```

This shows that for every year increase in age, the chance of not being married decreases by a factor of .8851. 

#### Conclusion

Overall, I liked how this logistical regression turned out. I think that it was interesting to see the probability of not being married at different ages and finding that age is definitely a factor in determining the likelihood of not being married once during a persons life. 
