---
title: "Day Care Fines"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r include=FALSE}
library(mosaic)
library(DT)
library(pander)
library(car)
library(tidyverse)
library(lattice)

# To get the "late" dataset into your R Console, 
# you need to go to your file menu of R-Studio and select
# "Session, Set working directory, To source file location"
# Then click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 
late <- read.csv("../Data/late.csv", header=TRUE)
# Then type:
# > View(late)
# in your R Console
```



----

<!--If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->

**Comments to Critiquers:** 

Thanks for the critique!

<!--End your comments to your critiquers before this line.-->

----
# {.tabset}

## Overview
<!-- Be sure to add your own background here before beginning the Analysis. -->

This overview is providing a quick summary of the detailed explanation of the study that is found under the "Study Details" tab. The study involves 10 different day care centers and roughly 35 children or less per day care center, but there are exceptions. The parents sometimes arrive late to pick up their children from the day care centers, and there were no fines for arriving late. At the beginning of the study, the parents were starting to be fined for showing up late to pick up their children. However, the parents were rarely more than 30 minutes late. The purpose of this test is to see if the fine (of roughly $3 per child) was enough to start changing the parents behavior and reduce the amount of parents that showed up late. 




<!--End your background before here.-->
<!--Click the little black arrow to 
the left to expand/hide these next sections.-->
## Study Details

<div style="color:#a8a8a8;">
This background is quoted directly from the article ["A Fine is a Price"](http://rady.ucsd.edu/faculty/directory/gneezy/pub/docs/fine.pdf). 
</div>

<div style="padding-left:30px; padding-right:30px;">

There are two types of day-care centers in Israel: private and public. A study was conducted in 10 private day-care centers in the city of Haifa from January to June 1998. All of these centers are located in the same part of town, and there is no important difference among them. During the day children are organized into groups according to age, from 1 to 4 years old. Each day-care center is allowed to hold a maximum of 35 children. In some exceptional cases a few additional children are allowed. The fee for each child is NIS 1,400 per month. (The NIS is the New Israeli Shekel.) At the time of the study, a U.S. dollar was worth approximately NIS 3.68, so the fee was about $380 at that time. 

The contract signed at the beginning of the year states that the day-care
center operates between 0730 and 1600. There is no mention of what happens if parents come late to pick up their children. In particular, before the beginning of the study, there was no fine for coming late. When parents did not come on time, one of the teachers had to wait with the children concerned. Teachers would rotate in this task, which is considered part of the job of a teacher, a fact that is clearly explained when a teacher is hired. Parents rarely came after 1630. 

A natural option [to fix the problem of parents showing up late] is to introduce a fine: every time a parent comes late, [they] will have to pay a fine. Will that reduce the number of parents who come late? If the fine is removed, will things revert back to the way they were originally?

The overall period of the study was 20 weeks. In the first 4 weeks we
simply recorded the number of parents who arrived late each week. At the
beginning of the fifth week, we introduced a fine in six of the 10 day-care centers, which had been selected randomly. The announcement of the fine was made with a note posted on the bulletin board of the day-care center. Parents tend to look at this board every day, since important announcements are posted there. The announcement specified that the fine would be NIS 10 for a delay of 10 minutes or more. The fine was per child; thus, if parents had two children in the center and they came late, they had to pay NIS 20. Payment was made to the principal of the day-care center at the end of the month. Since monthly payments are made to the owner during the year, the fines were added to those amounts. The money was paid to the owner, rather then to the teacher who was staying late (and did not get any additional money). The teachers were informed of the fine but not of the study. Registering the names of parents who came late was a common practice in any case.

At the beginning of the seventeenth week, the fine was removed with no explanation. Notice of the cancellation was posted on the board. If parents asked why the fines were removed, the principals were instructed to reply that the fine had been a trial for a limited time and that the results of this trial were now being evaluated.

A comparison with other fines in Israel may give an idea of the size of
the penalty that was introduced. A fine of NIS 10 is relatively small but not insignificant. In comparison, the fine for illegal parking is NIS 75; the fine for driving through a red light is NIS 1,000 plus penalties; the fine for not collecting the droppings of a dog is NIS 360. For many of these violations, however, detection and enforcement are low or, as in the case of dog dirt, nonexistent in practice. A baby-sitter earns between NIS 15 and NIS 20 per hour. The average gross salary per month in Israel at the time of the study was NIS 5,595.

</div>


## The Data (Wide)


The `late` Day Care Center data is shown here in the "wide data format". 

```{r, out.width=300}
#Show the full width of the "Wide" version of the late data:
pander(late, split.tables = Inf)
```


## The Data (Long)

The `Late` Day Care Center data is shown here in the "long data format".

```{r}
# This code reshapes the data into "long" format called Late.
# To get the "Late" dataset into your R Console, 
# you need to click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 
# Then type:
#   > View(Late)
# in your R Console
Late <- reshape(late,
                varying = paste("Week",1:20, sep=""), 
                v.names = "No.ofLateChildren",
                timevar = "Week", 
                times = 1:20, 
                idvar = "Center",
                new.row.names = 1:200,
                direction = "long")
pander(Late)

```


 

<!-- You can begin writing again after here.-->

## Analysis

### Questions and Hypotheses

Question 1: Does the type of treatment affect the average number of late children?

Hypothesis 1: 

$$H_o: \mu_C = \mu_F = \mu$$
$$H_a: \mu_i ≠ \mu \ for \ at \ least \ one \ i \ \epsilon \left\{  1=Control,\ 2 = Fine\right\} $$
Question 2: Does the type of stage affect the average number of late children?

Hypothesis 2: 
$$H_o: \mu_A = \mu_B = \mu_C = \mu$$
$$H_a: \mu_i ≠ \mu \ for \ at \ least \ one \ i \ \epsilon \left\{  1=A,\ 2 = B, \ 3=C\right\} $$
(In this hypothesis, A = 'Pre-Experiment', B = 'Experiment', C = 'Post-Experiment')


Question 3: Does the effect of the stage of the experiement change for different types of treatment?

Hypothesis 3:

$$H_o: The \ effect \ of \ the \ stage \ of \ the \ experiment \ is \ the \ same \ for \ all \ types \ of \ treatment $$
$$H_a: The \ effect \ of \ the \ stage \ of \ the \ experiment \ is \ not \ the \ same \ for \ all \ types \ of \ treatment $$

The significance level for this test will be set to $\alpha = .05$

### ANOVA Test

```{r echo=TRUE}

Late <- Late %>% 
  mutate(
    Stage = case_when(
      Week == 1| Week == 2 | Week == 3 | Week == 4 ~ "Pre-Experiment",
      Week == 5| Week == 6 | Week == 7 | Week == 8 | Week == 9| Week == 10 | Week == 11 | Week == 12 |       Week == 13 | Week == 14 | Week == 15 | Week == 16 ~ "Experiment",
      Week == 17 | Week == 18 | Week == 19 | Week == 20 ~ "Post-Experiment"
    )
  )  

myaov <- aov(No.ofLateChildren ~ Treatment + as.factor(Stage) + Treatment:as.factor(Stage), data = Late)
summary(myaov)
```

The ANOVA test above shows that all three hypothesis have significant results, or in other words, the p-value is below the significance level. This means that for all three tests we can reject the null hypothesis. The meaning of this finding will be explained in more detail below. 

### Confirming the validity of the ANOVA results

```{r echo = TRUE}

plot(myaov,which = 1:2)
```

The two graphs above help to prove the validity of the results. For this ANOVA test to be valid, the variance needs to be constant and the data needs to be distributed normally. Based on the Residuals vs Fitted graph, the data can assumed to be constant and based on the Normal Q-Q graph, the data can assumed to be normal. 

### Graphs

```{r}
xyplot(No.ofLateChildren ~ Treatment, data=Late, type=c("p","a"), main="Significance of Treatment", col='gray') 
```

### Interpretation

The above graph shows an upward slanting line which shows that there is a difference in means for the results between Control and Fine. I thought it was interesting that the children whose parents were fined actually had more late children than the control group. 

```{r}
xyplot(No.ofLateChildren ~ relevel(as.factor(Late$Stage), ref = 3), data=Late, type=c("p","a"), main="Significance of Stage", xlab = "Stage", ylab = "No. of Late Children")
```

### Interpretation

Based on the results above, this shows that between the pre-experiment and the experiment itself, there was an increase in the number of late children. However, after the fine was lifted, this average remained the same. I would have thought it would have gone down. 

```{r}
xyplot(No.ofLateChildren ~ relevel(as.factor(Late$Stage), ref = 3), data=Late, groups = Treatment, type=c("p","a"), main="Significance of the Interaction", auto.key=list(corner=c(1,1)), xlab = "Stage", ylab = "No. of Late Children")
```

### Interpretation

Lastly, this final graph shows that the control remained mainly unaffected throughout the experiment, which is in line with expectations, since it is the control group I would expect it to remain consistent. The fine group showed that there was interaction between the Pre-Experiment and the Experiment. 

### Conclusion 

I did not expect that the results of imposing a fine would make more parents show up late than before. This could be due to the fact that some parents feel that if they pay the fine then it is okay to show up late. It could also be because the fine was not high enough to cause parents to try avoid showing up late. It was also interesting to see that after this initial increase of parents showing up late from the experiment, it did not decrease after the experiment ended. This result could be because showing up late became a habit for some of the parents. If I was consulting the day care centers, I would tell them to not impose a fine if they wanted more parents to show up on time. 
