---
title: "Recalling Words"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(car)
library(pander)
library(DT) #You may need to run: install.packages("DT") 
```

```{r, eval=FALSE}
# Play the chunk above and this one to get the data into your Console
View(Friendly)
?Friendly
```
<br />

# {.tabset}

## Background

Many teachers and other educators are interested in understanding how to best deliver new content to students. In general, they have two choices of how to do this.

1. The Meshed Approach
    * Deliver new content while simultaneously reviewing previously understood content.

2. The Before Approach
    * Deliver new content after fully reviewing previously understood content.

A study was performed to determine whether the *Meshed* or *Before* approaches to delivering content had any positive benefits on memory recall. 

<div style="padding-left:15px;">

##### <a href="javascript:showhide('uniquename')">The Experiment <span style="font-size:8pt;">(click to view)</span></a>


<div id="uniquename" style="display:none;">

Individuals were seated at a computer and shown a list of words. Words appeared on the screen one at a time, for two seconds each, until all words had been shown (40 total). After all words were shown, they were required to perform a few two-digit mathematical additions (like 15 + 25) for 15 seconds to avoid immediate memory recall of the words. They were then asked to write down as many of the 40 words as they could remember. They were given a maximum of 5.3 minutes to recall words.

The process of showing words and recalling words was repeated four times with the same list of words each time (four chances to get it right). The presentation of the first trial was the same for all treatment conditions. However, trials 2, 3, and 4 were slightly different for each treatment condition.

<div style="padding-left:15px;">

The `SFR` group (the control group) stands for Standard Free Recall. In all four trials the same list of 40 words was presented, in a random order each time.

The `Before` group also used the same 40 words during each trial. However, any words that were correctly recalled in a previous trial were presented first, or *before* the words that were not recalled in the last trial. After all the correct words were presented in random order, the non-recalled words were presented in a random order.

The `Meshed` group also used the same 40 words during each trial. However, words that were correctly recalled in a previous trial were alternated with a missed word during the next presentation order. 

</div>

The data records the number of correctly recalled words (out of the 40 possible) from the fourth trial. Results were obtained for 30 students, 10 in each of the three treatment groups: `SFR`, `Before`, and `Meshed`. 

</div>

##### <a href="javascript:showhide('uniquename2')">The Data <span style="font-size:8pt;">(click to view)</span></a>

<div id="uniquename2" style="display:none;">

The results from the study can be found in the `Friendly` data set in R after loading `library(car)`. 

Click the "Code" button to see the data.


```{r}
datatable(Friendly, options=list(lengthMenu = c(3,10,30)))
```


</div>
</div>

<br />

```{r include=FALSE, echo=FALSE}
library(car)
library(tidyverse)
library(graphics)
```

```{r echo=TRUE}
dat <- Friendly
datSFR <- subset(dat, condition == "SFR")
datBefore <- subset(dat, condition == "Before")
datMeshed <- subset(dat, condition == "Meshed")
```

## Question and Hypothesis

The question that the following tests are trying to answer is, "Which of the three methods (SFR, Before, or Meshed) mentioned are most effective in helping a student memorize information?" This is intriguing since proving that one method is more affective than another would influence teaching styles to help students. My hypothesis is that the Before group will perform the best when compared with the other groups, which means that I believe that the three methods will not all be stochastically equal. 

$$ H_o: the\ distributions\ are\ stochastically\ equal $$
$$ H_a: the\ Before\ group\ is\ stochastically\ greater\ than\ the\ SFR\ and\ the\ Meshed\ groups $$


I will test my hypothesis by performing a wilcoxon rank sum test. I will also display the results through the use of box plots. 

##My Tests and Visualizations

#### Testing the data for normal distribution
```{r echo=TRUE}
par(mfrow=c(1,3))
qqPlot(datSFR$correct, main = "SFR Group", ylab = "Correct Number of Words Recalled")
qqPlot(datBefore$correct, main = "Before Group", ylab = "Correct Number of Words Recalled")
qqPlot(datMeshed$correct, main = "Meshed Group", ylab = "Correct Number of Words Recalled")
```

####Explanation

These qqplot graphs were made in order to test if the data was normally distributed. I split up the test into three groups: the SFR group, the Before group, and the after group. It seems like the SFR group had normally distributed data, but the Before and Meshed Group both had outliers and therefore it would be more appropriate to do a Wilcoxon rank sum test. Another reason to do the Wilcoxon rank sum test is because the three groups are independent from each other, otherwise the Wilcoxon rank-sign test would be more appropriate. 

#### Testing to see which method might be most effective

```{r}
wilcox.test(datBefore$correct,datSFR$correct,mu= 0,alternative = "greater", conf.level = .975)
wilcox.test(datBefore$correct,datMeshed$correct,mu=0,alternative="greater", conf.level = .975)
```

#### Explanation

When comparing the SFR group and the before group, we can reject the null, but with the Before group and the Meshed group, we fail to reject the null.

#### Box Plots to further illustrate the wilcox test results

```{r echo=TRUE}
boxplot(dat$correct ~ dat$condition, col = c("blue","red","green"), ylab="Correct Number of Words Recalled")
```

## Conclusion

Overall, I found sufficient evidence to reject the null hypothesis that the Before group was stochastically greater than the SFR group. The reason I could come to this conclusion was because the alpha was set at .025 and the p-value was .022. However, evidence was insufficient to say that the Before group was stochastically greater than the Meshed group, therefore I failed to reject the null hypothesis in that regard. One final test that I want to do is see whether the Meshed group is stochastically greater than the SFR group as well. 

```{r echo=TRUE}
wilcox.test(datMeshed$correct, datSFR$correct, mu = 0,alternative = "greater", conf.level =.975)
```

The above results show that the Meshed group is not stochastically greater than the SFR group. 

The reason I set the alpha levels at .025 was because the original test was to see if the Before group was stochastically greater than both the Meshed and the SFR groups and since I am making two comparisons, I wanted to lower the alpha levels so the overall alpha level is near .05. In conclusion, the Before group was not stochastically greater than both the Meshed and SFR groups and I failed to reject the null hypothesis overall. 


